Practical Security Log Book

### 2014-12-15 logbook.tex
Finished writing lab 3.

IP fragmentation \parencite{postel1981rfc,fall2011tcp} is necessary if a router
receives a packet larger than the maximum transmission unit (MTU) of the
next hop. When this occurs, a router has the option to break the datagram
into smaller fragments. RFC 815 \parencite{clark1982ip} describes
some algorithms for how downstream hops can then reassemble the fragmented
IP packets.

@@

In this report, we look at how to set up GNU Privacy Guard (GPG)
on a Debian or Ubuntu GNU/Linux
system although many of the steps apply to other Linux and Unix systems.

The steps outlined below will demonstrate how to:

\begin{enumerate}
\item generate a new private/public GPG key pair;
\item have a someone sign your new public key to verify that you own it;
\item use the public key to encrypt a plain text message; and
\item use the private key to descrypt the message.
\end{enumerate}

@@

The gpg tool is normally pre-installed in a lot of Linux/Unix systems as
tools like \texttt{apt} depends on it. On
a Debian or Ubuntu GNU/Linux system that does not have it installed, it
can be installed quite simply with:

@@

Note that on the first use of the command, it creates some configuration files
and keyrings. If there are any keys already stored in the user's home
directory, then details of them would be printed out as well.

@@

After answering all the questions and providing a passphrase,
you may be asked to generate some entropy.
Machines had a hard time using a good source of randomness to make sure
the key is not predictable and guessable, so you may be required to do
some work via keystrokes, mouse movements or run some processes that
do disk activity. Once this is complete, you will have your public and
private key pair.

@@

they are then able to encrypt this secret message so that only you may read it:

@@

Now we have the original message intact, but also an encrypted form of it.
Again, the \texttt{--armor} option would have created an ASCII form of
the encrypted message more suitable for pasting into emails (in a file
called secret.txt.asc).

@@

\chapter{Ethical Hacking Lab Setup}

@@

>>>>

IP fragmentation \parencite{postel1981rfc,fall2011tcp} is necessary if a router
receives a packet larger than the maximum transmission unit (MTU) of the
next hop. When this occurs, a router has the option to break the datagram
into smaller fragments. RFC 815 \parencite{clark1982ip} describes
an algorithm for how downstream hops can then reassemble the fragmented
IP packets.

@@

In this report, we look at how to set up GNU Privacy Guard (GPG)
on a Debian or Ubuntu GNU/Linux
system although many of the steps apply to other Linux and Unix systems.
The steps outlined below will demonstrate how to:

\begin{enumerate}
\item generate a new private/public GPG key pair;
\item have a someone sign your new public key to verify that you own it;
\item use the public key to encrypt a plain text message; and
\item use the private key to decrypt the message.
\end{enumerate}

@@

The gpg tool is normally pre-installed in a lot of Linux/Unix systems as
tools like \texttt{apt} depend on it. On
a Debian or Ubuntu GNU/Linux system that does not have it installed, it
can be installed quite simply with:

@@

Note that on the first use of the command, it creates some configuration files
and keyrings. If there were any keys already stored in the user's home
directory, then details of them would be printed out as well.

@@

After answering all the questions and providing a passphrase,
you may be asked to generate some entropy.
Machines have a hard time using a good source of randomness to make sure
the key is not predictable and guessable, so you may be required to do
some work via keystrokes, mouse movements or run some processes that
do disk activity. Once this is complete, you will have your public and
private key pair.

@@

\noindent they are then able to encrypt this secret message so that only you may read it:

@@

Now we have the original message intact, but also an encrypted form of it.
Again, the \texttt{--armor} option would have created an ASCII form of
the encrypted message more suitable for pasting into emails (in a file
called \texttt{secret.txt.asc}).

@@

\chapter{Ethical Cracking Lab Setup}

In this report, we look at some tools that can be used to put together
a small lab setup for practising computer cracking and penetration
testing. This setup could also be used to observe and trace behaviours
of malware, e.g. viruses, but perhaps with a more stringent approach
to network isolation (see section~\ref{sec:netsec}).

The setup here is a network of virtual machines running under Virtualbox
running on a host computer that can be isolated if necessary. Once the
setup has been designed, we will look at using Vagrant to make a repeatable
way of bootstrapping the lab on demand. This will help if we ever feel the
need to tear down the entire setup and start afresh or to share the
configuration with other researchers.

\section{Kali}

Kali is a GNU/Linux distribution based on Debian that comes with hundreds
of tools preinstalled that can be used in cracking and penetration testing. It
is available to download as an ISO CD image file which can easily be used
to install in a virtual machine using Virtualbox.

The Vagrantfile configuration in appendix~\ref{appendix:vagrant} shows a
simple provision of a Virtualbox instance of Kali from a pre-built
Vagrant box for convenience. For more control or more assurance about the
security of the virtual machine, it should be possible to fork and customise
the Packer build that created this image.\footnote{https://github.com/ctarwater/packer-kali}

\section{Metasploitable}

Metasploitable is a pre-built virtual machine from Rapid 7 to aid with learning
its penetration testing tool, Metasploit. We can equally make use of this
for learning Kali. As with Kali, we can add a configuration to our Vagrantfile
(appendix~\ref{appendix:vagrant}) for a pre-built box for convenience. Other
Packer-based builds are available on Github\footnote{https://github.com/waratek/vagrant-boxes}
if it is again preferred to have more control over the build.

\section{Vagrant}

As discussed in the previous two sections, vagrant provides us two conveniences
in building our lab setup:

\begin{enumerate}
\item Repeatable, refreshable and shareable builds of our whole lap setup. A whole classroom
  of lab setups for teaching could be running within an hour.
\item The ability to use pre-built box images, at least in the first instance, to get set up
  quickly, but still migrate to in-house builds of those images when the time is right.
\end{enumerate}

We can configure vagrant with the \emph{Vagrantfile} shown in appendix~\ref{appendix:vagrant},
which is a reasonably trivial configuration to bootstrap both Kali and Metasploitable on
the same private network. Note that the GUI is enabled so we can log in to X or a TTY
on either machine, which is required because the NAT adapter is overridden such that the
machines cannot communicate with the outside world. A side-effect of this is that
vagrant is unable to configure the network interfaces on the guest machines automatically
for us and we must edit \texttt{/etc/network/interfaces} ourselves, e.g.:

\begin{lstlisting}
iface eth0 inet static
    address 192.168.51.5
    netmask 255.255.255.0
\end{lstlisting}

The reasoning around this is discussed in section~\ref{sec:netsec}.

\section{Network Security}
\label{sec:netsec}

By overriding the NAT interface vagrant wants to use, we can enforce that
our lab network is only using a host-based adapter on a private network such
that the guests do not even have a gateway to the outside world. We can see
on the routing table that the guest only knows how to talk to the private
network:

\begin{lstlisting}[language=bash]
vagrant@metasploitable:~$ route
Kernel IP routing table
Destination     Gateway         Genmask         Flags Metric Ref    Use Iface
192.168.51.0    *               255.255.255.0   U     0      0        0 eth0
vagrant@metasploitable:~$
\end{lstlisting}

For further security, it may be desirable to add iptables rules on the host
machine to drop any traffic that isn't going between the two defined
static IPs. This is perhaps an unnecessary step if we are only using the
penetration and exploit testing tools by hand and are in control of what
we are doing.

This lab setup can however be extended to include Windows
machines in the Vagrantfile such that we are able to examine malware and
viruses. If we are running malicious code (as opposed to simply running
exploitable services), then we likely want to make sure the host machine
is simply disconnected from any network and reinstall the whole host
before it is allowed back on any public network.

@@

\chapter{Ethical Cracking Lab Vagrant Configuration}
\label{appendix:vagrant}

\lstinputlisting[basicstyle=\scriptsize,language=ruby]{networks/pentest/Vagrantfile}
